\bf Approachable case studies support learning and reproducibility in data science: An example from evolutionary biology

### 2022-02-16 jsdse/Manuscript.tex
almost there

\if0\blind
{
  \title{\bf Approachable case studies support learning and reproducibility in data science: An example from evolutionary biology}
  \author{Luna L. Sanchez Reyes\thanks{
    The authors gratefully acknowledge ``Sustaining the Open Tree of Life``, NSF ABI No. 1759838, and ABI No. 1759846.}\hspace{.2cm}\\
    School of Natural Sciences, University of California, Merced\\
    and \\
    Emily Jane McTavish \\
    School of Natural Sciences, University of California, Merced}
  \maketitle
} \fi

@@

Using an example within evolutionary biology, we identify aspects of research workflows
that make them unapproachable to the general audience:
use of specialized computing language and programming techniques;
high cognitive load;
unspecified, unclear or lengthy goals;
content-focused descriptions instead of user-focused;
inflexible learning environment;
unapproachable (cold or intimidating) language; and
little to no diversity of representation of information.
Then, we propose a set of principles to improve the unapproachable aspects of
research workflows, and illustrate their application in a case study from
evolutionary biology that we used as teaching material.
Finally, we elaborate on the general application of these principles for
documenting research workflows and products, to provide present learners
and future researchers with tools for successful scientific reproducibility.
%> ** add a sentence on implications for teaching reproducibility

@@

\noindent%
{\it Keywords:}  open science, R, phylogenetics, Open Tree of Life
\vfill

@@

In this paper, we argue that research products that are digitally available are
often unapproachable in practice (Box 1), because they are not
friendly for learners and interested parties with different levels of
experience in the field. Research products that are unapproachable
counteract availability, and hinder reproducibility short and long term.
To support long term adoption of reproducible practices in the natural sciences, research
workflows need to be made approachable for learners, the researchers of the future.

@@

\bigskip
\bigskip

@@

Phylogenetics is a key discipline within evolutionary biology \citep{dobzhansky1973nothing}.
It focuses on investigating the history of shared ancestry of living and extinct
organisms using biological data,
and represents this evolutionary history with a diagram known as a phylogeny
or phylogenetic tree (because it grows through time and appears to have branches;
Figure \ref{fig:tree}).
Phylogenies provide the basis to study and understand all biological processes
in an evolutionary context \citep{dobzhansky1973nothing}
Hence, it appears that improving reproducibility rates in phylogenetics has the
potential to positively impact research across the natural sciences.

To explore barriers to approachability in phylogenetics, we develop a case study
that touches on three common problems within the field: standardizing
organism names in phylogenies, obtaining current phylogenetic knowledge for a group of organisms,
and summarizing this phylogenetic knowledge in a meaningful way.
To address these problems, we propose a research workflow that relies on resources from the Open
Tree of Life (OpenTree), an open source project that provides
digital availability of phylogenetic results from published, peer-reviewed research, which
is considered as vetted and state-of-the-art knowledge in the field.
OpenTree phylogenies are stored in a public database, the Phylesystem \citep{mctavish2015phylesystem},
and are downloadable as various computer-readable file types, which is key for reusability
and reproducible workflows \citep{wilson2017good}.
OpenTree also provides access to a single naming standard for organisms (taxonomic standard)
that is applied to the stored phylogenies \citep{rees2017automated}, which are
then used to summarize a single phylogenetic tree encompassing all life \citep{opentreeoflife2019synth}.

All of OpenTree resources
%% (the stored phylogenies, the taxonomic standard, the synthetic tree, and all
%% tools developed to process and summarize stored phylogenies),
are free of cost to any user, and are available for download and use through its Graphical User
Interface (GUI; aka, a website or application that allows users to access computer functionalities, in this case OpenTree resources, with mouse or keyboard clicks).
However, reducing as many manual steps as possible in research workflows is key
for reproducibility, as manual data manipulation scales poorly and is prone to
error \citep{bakken2019journey}.
OpenTree's resources are also programmatically available through its Application
Programming Interface services (APIs; aka, computer code that implements computer functionalities,
in this case OpenTree resources, that can be used by programmers
to build more functionalities),
which provide scalability and reproducibility \citep{opentreeAPIv3}.
However, this comes at a high cost for the user, which requires considerable more
computer programming experience and literacy to be able to successfully use APIs.
%% In recent years, OpenTree API services have been wrapped into software packages for R and Python
%% \citep{michonneau2016rotl, mctavish2021opentree}, open source, free of cost
%% programming languages that represent two of the most widely used programming languages
%% in the sciences \citep{baker2017scientific}. These packages should contribute to making
%% OpenTree resources more accessible to a wider user audience.
The \texttt{rotl} R package \citep{michonneau2016rotl} and the \texttt{opentree} Python module
\citep{mctavish2021opentree} have been developed as wrappers for OpenTree's API services.
R and Python are open source and free of cost programming languages that represent
two of the most widely used programming languages in the sciences today \citep{eglen2009quick, baker2017scientific}.
As such, \texttt{rotl} and \texttt{opentree} software packages should contribute
to making OpenTree resources more accessible to a wider programming user audience.

@@

In the next section, we describe the barriers to approachable research workflows that we identified
on our case study. We address these barriers in a set of teaching materials that are
available at
https://mctavishlab.github.io/R\_OpenTree\_tutorials/.

@@

The main goal of our research workflow is to obtain a single phylogeny summarizing
data from a set of published phylogenies for the canids (the family of dogs, coyotes, wolves, etc.),
our organism of study.
Our analysis can be completely accomplished using functions
from the R package \texttt{rotl} or the Python module \texttt{opentree}. If
a researcher were to use this workflow in a publication, they would typically
describe it in the methodology section as ``The canid summary phylogeny was obtained using
functions from X package, details are available as supplementary files".
This is usual practice, mainly because journals do not have space to publish
all code used in an analysis in the methods section. Yet, supplementary files have
the misfortune to not be peer-reviewed as thoroughly (or at all) as the main manuscript
\citep{CITATION}.
They are also prone to the dreaded promise ``available upon request", which has
very low rates of fulfillment \citep{CITATION}.
Without the primary code that was used to perform an analysis it is impossible
to reproduce said analysis. When the code is available, other issues can
complicate reproduction of the analysis to the point of completely obstructing
reproducibility.
For example, what software do I need to open the code files? Can I run the code within the same
software, or do I need a different one?
Do I need additional software that the analysis depends on? What does the code even mean?

These questions are usually addressed in the software documentation.
As opposed to code, documentation is written in natural language (i.e.,
any known human language, e.g., English, Spanish, Chinese), and is considered
a key element for successful adoption of software
by the target users \citep{karimzadeh2018top}. This might explain
why documentation for software addressed to academic users is also
usually written using highly specialized computational
language or jargon (i.e., computationally specific concepts,
words, and phrases) as well as formal scientific and academic language.
\textit{Barrier 1.--} While scientific jargon might have an important role for formal acceptance
of software by the scientific and academic community, it can be perceived as
cold and/or intimidating language, that often slows down or even
obstructs examination, application, and adoption of code by a wider audience \citep{ball2017its},
and discourages learners by creating a hostile environment that does not foster learning \citep{CITATION}.

Good software documentation for code has to be thorough \citep{karimzadeh2018top}.
It should describe general usage of individual functions, as well as
the argument and variables said function can take, and it should be accompanied with
usage examples on how to apply each function \citep{karimzadeh2018top}.
Individual documentation for each function is usually presented
in alphabetic order, and does not have a specific goal, besides being as thorough as possible.
In this context, identifying connections across functions that are meant to work on the same
analysis workflow is difficult. Moreover, most software has many functions, so documentation
is usually very lengthy. \textit{Barrier 2.--}  All of this increases the cognitive
load for the target users and learners, which discourages learners \citep{CITATION}.

Another important aspect of software documentation is that examples are usually
worked and showcase the ideal or minimal case in which a function should work well.
\textit{Barrier 3.--} Focusing importance on the content of the software instead
of considering the user experience, ignoring what can go wrong and avoiding clear
advice on how to troubleshoot.

Software documentation is usually explained with words, very rarely using diagrams or
other allegories.
\textit{Barrier 4.--} Little to no diversity of representation of information

Documentation should always be available for the users and learners
\textit{Barrier 5.--} Inflexible learning environment, is this one really an issue???

% Some principles of an inclusive syllabus:

% - unspecified, unclear or lengthy goals; high cognitive load; (literate programming)

% - content-focused descriptions/examples instead of user-focused; (demonstrate errors)

% - inflexible learning environment; (make it stable so users can learn whenever is best for them)

% - unapproachable (cold or intimidating) language, use of highly specialized computing jargon, skill level; (use friendly and informal but respectful language)

In sum, best practices for good primary documentation are not enough to ensure
reproducibility of research workflows that rely heavily on code.

\section*{Best practices for approachable research workflows}
\label{sec:addressing}

\subsubsection*{a. Reduce cognitive load and provide specific and clear goals
with literate programming}

Pedagogical research shows that active learning practices are one the most effective
ways to take on abstract subjects \citep{freeman2014active}.
Programming computer languages are quite abstract and cognitive load can be greatly
reduced for learners by applying an active learning strategy such as linking its usage to
a ``real world'' or ``human'' application \citep{felder2009active}.

@@

\subsubsection*{b. Provide examples that are user-focused by demonstrating errors and warnings}

@@

A practice that has become more and more widespread in programming-language pedagogical practices
is the use of typos and mistakes to normalize them for learners, and show them how
to solve them when they are outside the classroom \citep{shannon2015live}.
Yet, this is rarely done on written pedagogical materials.
Primary documentation focuses on demonstrating usage function with examples that
work seamlessly, without errors. We argue that the opposite is needed to support
adoption of reproducible workflows and support long term independence in learners
\citep{gaspar2007restoring}.
We demonstrate examples that do not work
as expected and exemplify ways to address them (Figure \ref{fig:error}).

We identify inputs that would give
a wide range of warnings and errors, focusing on demonstrating these cases. This
helps users to not be afraid of errors and warnings, but instead to use them to
their advantage.
We also identify effects of warnings and errors downstream of the workflow.

We identify ways to evaluate inputs to know if they will produce an error, and design
alternatives on what to do when faced with an error or warning, and demonstrate
these alternatives.
One of the most essential skills in programming is interpreting and moving forward
from errors.
Many finely honed tutorials do not trigger errors, which precludes helping students
to develop the tools to understand and address errors when they do encounter them,
as they inevitably will.
In our tutorial, we focus on explaining the meaning and downstream of warnings and errors, and
 showcase ways to detect them before they are triggered (i.e., before using an input
  that would elicit a warning or error). This has two pedagogical benefits:
1) it provides users/students with the means to troubleshoot their own warnings and errors, and
2) it allows users/students to understand with more depth what the function is doing.

% We designed ways to access the different elements of the outputs.

\subsubsection*{c. Use friendly, relatable and respectful language}

\subsubsection*{d. Make it accessible geographically and through time}

% - inflexible learning environment; (make it stable so users can learn whenever is best for them)

We published the tutorials on a public, free license, free of cost, and free for
use and reuse repository and persistent website \citep{RopentreeTutorials}.
The tutorial is available for the users to go back to any time they need it,
and to be passed on to other users (Figure \ref{fig:schedule}).

\begin{figure}
\begin{center}
\includegraphics[width=3in]{fig-schedule.png}
\end{center}
\caption{Snapshot of the home to our tutorial website, showing part of the schedule. \label{fig:schedule}}
\end{figure}

We created a main version of the tutorial that is stable. Any updates to the tutorial are
published as new versions, or tutorials for new workshops \citep{wilson2006swc, SWCwebsite}
 Versions presented at workshops are a copy from the original repository.
They represent a temporally stable snapshot of functions and workflows presented
during a workshop.

%%% \subsubsection*{5. The (now) classics of computational reproducibility}
%%%
%%% Provide all information on package version and system capabilities.
%%%
%%% > [name=Emily Jane McTavish]link to where/and how you did that

@@

Response form the community has been invaluable in gauging success of our teaching
materials.
Senior researchers often comment on the usefulness
of the tutorials for their research, as well as how they have supported students in using the
R packages with less help from them as PIs.

Making accessible reproducible workflows has several advantages:

@@

Ultimately, the long term improvement of reproducibility rates in science will depend
on our ability to intentionally integrate the subject of reproducibility into the
undergraduate curriculum, so college learners and future researchers have the
basis to develop the fundamental skills needed to successfully create reproducible
scientific workflows and materials.

@@

>>>>

\if0\blind
{
  \title{\bf Approachable case studies support learning and reproducibility in data science: An example from evolutionary biology}
  \author{Luna L. Sanchez Reyes\thanks{
    The authors gratefully acknowledge ``Sustaining the Open Tree of Life'',
    NSF ABI No. 1759838, and ABI No. 1759846. They also acknowledge ``The Carpentries''
    project, without it this project could not have been possible}\hspace{.2cm}\\
    School of Natural Sciences, University of California, Merced\\
    and \\
    Emily Jane McTavish \\
    School of Natural Sciences, University of California, Merced}
  \maketitle
} \fi

@@

Using a case study within evolutionary biology, we identify aspects of research workflows
that make them unapproachable to the general audience:
use of highly specialized language is intimidating;
unspecified, unclear or lengthy goals lead to high cognitive load;
content-focused descriptions instead of user-focused; and
little to no diversity of representation of information.
Then, we propose a set of principles to improve the unapproachable aspects of
research workflows, and illustrate their application in a case study from
evolutionary biology that we used as teaching material.
Finally, we elaborate on the general application of these principles for
documenting research workflows and products, to provide present learners
and future researchers with tools for successful scientific reproducibility.
%> ** add a sentence on implications for teaching reproducibility

@@

\noindent%
{\it Keywords:}  open science, R, phylogenetics, Open Tree of Life, pedagogy
\vfill

@@

In this paper, we argue that research products that are digitally available are
often unapproachable in practice, because they are not
friendly for learners and interested parties with different levels of
experience in the field. Research products that are unapproachable
counteract availability, and hinder reproducibility short and long term.
To support long term adoption of reproducible practices in the natural sciences, research
workflows need to be made approachable for learners, the researchers of the future \citep{roland2002think}.

@@

@@

Phylogenetics is a key discipline within evolutionary biology \citep{dobzhansky1973nothing}.
It focuses on investigating the history of shared ancestry of living and extinct
organisms using biological data,
and represents this evolutionary history with a diagram known as a phylogeny
or phylogenetic tree (because it grows through time and appears to have branches;
Figure \ref{fig:tree}).
Phylogenies provide the basis to study and understand all biological processes
in an evolutionary context \citep{dobzhansky1973nothing}.
Hence, it appears that improving reproducibility rates in phylogenetics has the
potential to positively impact research across the natural sciences.

To explore barriers to approachable phylogenetics, we develop a case study
that touches on three common problems within the field: standardizing
organism names in phylogenies, obtaining current phylogenetic knowledge for a group of organisms,
and summarizing this phylogenetic knowledge in a meaningful way.
To address these problems, we propose a research workflow that relies on resources from the Open
Tree of Life (OpenTree), an open source project that provides
digital availability of phylogenetic results from published, peer-reviewed research, which
is considered as vetted and state-of-the-art knowledge in the field.
OpenTree phylogenies are stored in a public database, the Phylesystem \citep{mctavish2015phylesystem},
and are downloadable as various computer-readable file types, which is key for reusability
and reproducible workflows \citep{wilson2017good}.
OpenTree also provides access to a single standard for organism names (taxonomic standard)
that is applied to the stored phylogenies \citep{rees2017automated}, which are
then used to summarize a single phylogenetic tree encompassing all life \citep{opentreeoflife2019synth}.

All of these resources
%% (the stored phylogenies, the taxonomic standard, the synthetic tree, and all
%% tools developed to process and summarize stored phylogenies),
are available for download and use from OpenTree free of cost to any user.
One way to access OpenTree resources is through it Graphical User
Interface (GUI; aka, a website or application that allows users to access computer
functionalities, in this case OpenTree resources, with mouse or keyboard clicks).
However, reducing as many manual steps as possible in research workflows is key
for reproducibility, as manual data manipulation scales poorly and is prone to
error \citep{bakken2019journey}.
OpenTree's resources are also programmatically available through its Application
Programming Interface services (APIs; aka, computer code that implements computer functionalities,
in this case OpenTree resources, that can be used by programmers
to build more functionalities),
which provide scalability and reproducibility \citep{opentreeAPIv3}.
However, this comes at a high cost for the user, which requires considerable more
computer programming experience and literacy to be able to successfully use APIs.
%% In recent years, OpenTree API services have been wrapped into software packages for R and Python
%% \citep{michonneau2016rotl, mctavish2021opentree}, open source, free of cost
%% programming languages that represent two of the most widely used programming languages
%% in the sciences \citep{baker2017scientific}. These packages should contribute to making
%% OpenTree resources more accessible to a wider user audience.
The \texttt{rotl} R package \citep{michonneau2016rotl} and the \texttt{opentree} Python module
\citep{mctavish2021opentree} have been developed to wrap OpenTree's API services
and make them available through the R and Python programming languages.
R and Python are open source and free of cost, and represent
two of the most widely used programming languages in the sciences today \citep{eglen2009quick,
baker2017scientific}.
As such, \texttt{rotl} and \texttt{opentree} software packages should contribute
to making OpenTree resources more accessible to a wider programming user audience.

@@

In the next section, we describe (in no particular order) the barriers to approachable research
workflows that we identified on our case study.
Then we develop a set of principles to address these barriers, and apply the latter
to a set of teaching materials that are available at
https://mctavishlab.github.io/R\_OpenTree\_tutorials/.

@@

The main goal of our case study is to obtain a single phylogeny summarizing
data from a set of published phylogenies for the canids (the family of dogs, coyotes, wolves, etc.),
our organism of study.
All analysis for our case study can be completely accomplished using functions
from the R package \texttt{rotl} or the Python module \texttt{opentree}. If
a researcher were to use the proposed analysis workflow in a publication, they would typically
describe it in the methodology section as ``The canid summary phylogeny was obtained using
functions from X package, details are available as supplementary files".
This is usual practice, mainly because journals do not have space to publish
all code used for an analysis in the methods section. Yet, supplementary files and data have
the misfortune to not be peer-reviewed as thoroughly (or at all) as the main manuscript
\citep{pop2015use}.
They are also prone to the dreaded promise ``available upon request", which has
very low rates of fulfillment \citep{krawczyk2012available}.
Without the primary data and code that was used to perform an analysis it is impossible
to reproduce said analysis \citep{miyakawa2020no}. When the code is available, other issues can
complicate reproduction of the analysis, to the point of completely obstructing
reproducibility.
For example, some questions that are often unaddressed in research workflows are:
what software do I need to open the code files?
Can I run the code with the same software I used to open the files, or do I need a different one?
Do I need additional software that the analysis depends on?
What does the code even mean?

These questions can usually be answered by referring to the software documentation,
which are usually publicly available and can be accessed by any potential user.
As opposed to code, software documentation is written in natural language (i.e.,
any known human language, e.g., English, Spanish, Chinese), and is considered
a key element for successful adoption of software
by target users \citep{karimzadeh2018top}. This might explain
why documentation for software addressed to academic users is also
usually written using highly specialized computational
language or jargon (i.e., computationally specific concepts,
words, and phrases) as well as formal scientific and academic language.
We identify this as \textit{barrier 1} to approachable research workflows.--
\textit{Specialized language is intimidating}.
While scientific jargon might have an important role for formal acceptance
of software by the scientific and academic community, it can be perceived as
cold and/or intimidating language, that often slows down or even
obstructs examination, application, and adoption of code by a wider audience
\citep{ball2017its}.
In contrast, introducing information without the use of jargon, supports learners'
conceptual understanding of new ideas and concepts \citep{mcdonnell2016concepts, pan2019online}.
% jargon creates a hostile environment that does not foster learning \citep{CITATION}.

Another element of good software documentation is that it has to be thorough \citep{karimzadeh2018top}.
Meaning that it should describe general usage of individual functions, as well as
arguments and variables that said function can take \citep{karimzadeh2018top}.
Individual documentation for each function is usually presented in alphabetic order,
and does not have a specific analysis goal.
Moreover, most software has numerous functions, so documentation is usually very
lengthy and it is hard to navigate.
This can have the effect of increasing the amount of information that needs to be
simultaneously processed by the users, which can lead to overload of the finite amount
of working memory any one possesses, known as cognitive load \citep{sweller1988cognitive}.
In this context, identifying connections across functions that are meant to work
on the same analysis workflow can become a very difficult task.
We recognize this as \textit{barrier 2.-- Lack of specific goals leading to high cognitive load}.
High cognitive load is know to have a negative effect in learning software
\citep{chandler1996cognitive, van2005research, lambert2009student}.

A third important aspect of software documentation are examples that demonstrate
usage of individual functions \citep{karimzadeh2018top}.
Examples presented in software documentation are usually
worked to perfection, as they are intended to showcase the ideal or minimal case
in which a function works well.
Perfectly worked examples ignore the user experience by
maintaining focus on the software content, and fail to provide users with expert
and clear advice on how to troubleshoot if needed.
We identify this as \textit{barrier 3.-- Content-focused examples}.
Error management training is an approach that focuses on framing mistakes as beneficial
to learning complex tasks, to give learners the opportunity to
actively explore a task \citep{frese1995error}.
Providing examples that showcase potential errors supports users performance \citep{steele2014error},
and can greatly improve learners ability to troubleshoot outside the classroom
\citep{shannon2015live, nederbragt2020ten}.

Last but not least, software documentation is usually explained verbally in a document, very
rarely is it approached using diagrams or other allegories.
We recognize this as \textit{barrier 4.-- Little to no diversity of representation of information}.

In sum, best practices for good software documentation are not enough to promote
reproducibility of research workflows that rely heavily on code. In the following
section we describe some principles that can help to reduce or remove the
identified barriers, to create
research workflows that are more approachable and hence more reproducible by a
larger audience.

\section*{Principles for approachable research workflows}
\label{sec:addressing}

\subsubsection*{Principle 1. Use friendly, relatable and respectful language}

\subsubsection*{Principle 2. Reduce cognitive load by providing specific and clear goals
with literate programming}

Cognitive load can be greatly reduced for learners by applying an active learning
strategy such as linking usage to a ``real world'' or ``human'' application
\citep{felder2009active}.
Programming computer languages are by themselves quite abstract and represent a
learning subject with a potentially high cognitive load for most learners.
Pedagogical research shows that active learning practices are one of the most effective
ways to take on abstract subjects \citep{freeman2014active}.

@@

\subsubsection*{Principle 3. Provide examples that are user-focused by demonstrating errors and warnings}

@@

A practice that has become more and more widespread in programming-language pedagogical practices
is live programming.
One benefit of live programming is that typos and mistakes occur -- which normalizes
them for learners, and shows them how to solve them when they are outside the
classroom \citep{shannon2015live, nederbragt2020ten}.
When co-author McTavish was a postdoc, teaching an introductory programming workshop as a volunteer
with the Carpentries (a non-profit group that teaches foundational coding and
 data science skills to researchers worldwide) \citep{wilson2006swc, SWCwebsite},
a senior faculty member taking the workshop
complained that the typos were slowing things down and interfering with the pedagogy.
McTavish replied "the typos ARE the pedagogy".
This has become a slogan of sorts at the Carpentries, capturing the idea that embracing
and discussing mistakes is essential to teaching programming \citep{wilson2019teaching}.
Yet, working through mistakes is rarely done on written pedagogical materials.
Software documentation focuses on demonstrating usage function with examples that
work seamlessly, without errors. We argue that the opposite is needed to support
adoption of reproducible workflows and support long term independence in learners
and users performance
\citep{gaspar2007restoring, steele2014error}.
In our tutorials, we apply this principle by demonstrating examples that do not work
as expected, and exemplifying ways to address them (Figure \ref{fig:error}).

We identified inputs that would give
a wide range of warnings and errors, and support learners on reading the information
provided by the message. This
helps users to not be afraid of errors and warnings, but instead use them as
an informative measure.
% This is probably part of the computer science curriculum, but it is not at all part of the natural sciences curriculum.
We also identify effects of warnings and errors downstream of the workflow.

We identify ways to evaluate inputs to know if they will produce an error, and design
alternatives on what to do when faced with an error or warning, and demonstrate
these alternatives.
One of the most essential skills in programming is interpreting and moving forward
from errors.

In our tutorial, we focus on explaining the meaning of warnings and errors, and
 showcase ways to detect them before they are triggered (i.e., before using an input
  that would elicit a warning or error). This has two pedagogical benefits:
1) it provides users/students with the means to troubleshoot their own warnings and errors, and
2) it allows users/students to understand with more depth what the function is doing.

\subsubsection*{Principle 4. Demonstrate one thing in different ways}

%We designed ways to access the different elements of the outputs.

@@

Response from the community has been invaluable in gauging success of our teaching
materials.
Senior researchers often comment on the usefulness
of the tutorials for their research, as well as how they have supported students in using the
R packages with less help from them as PIs.

One key aspect for success of our tutorials is that we applied all rules for
successful reproducibility, in particular making our tutorials available
through time in a public website.

Making approachable research workflows has several advantages:

@@

% \citep{roland2002think}
Ultimately, the long term improvement of reproducibility rates in science will depend
on our ability to intentionally integrate reproducibility into the
undergraduate curriculum, so college learners and future researchers have the
basis to develop the fundamental skills needed to successfully create reproducible
scientific workflows and materials.

@@

\bigskip
\bigskip
\bigskip

\begin{center}
{\large\bf FIGURES}
\end{center}

\begin{figure}
\begin{center}
\includegraphics[width=3in]{fig-schedule.png}
\end{center}
\caption{Snapshot of the home to our tutorial website, showing part of the schedule.
 Our tutorial website was constructed using The Software Carpentries workshop template \citep{swc-workshop-template}. \label{fig:schedule}}
\end{figure}
